### 2017-02-06 Documents/Euler/Benchmarks1D/draft.tex
fixes for compilation

\documentclass[10pt]{article}

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{booktabs}
\usepackage{mathrsfs}
\usepackage{cite}
%\usepackage{times}
\usepackage{url}
%\usepackage{hyperref}
\usepackage{natbib}
%\hypersetup{
 % bookmarksnumbered = true,
  %bookmarksopen=false,
  %pdfborder=0 0 0,         % make all links invisible, so the pdf looks good when printed
  %pdffitwindow=true,      % window fit to page when opened
  %pdfnewwindow=true, % links in new window
  %colorlinks=true,           % false: boxed links; true: colored links
  %linkcolor=blue,            % color of internal links
  %citecolor=magenta,    % color of links to bibliography
  %filecolor=magenta,     % color of file links
  %urlcolor=cyan              % color of external links
%}
\usepackage{graphicx}
\graphicspath{ {figures/} }

\newcommand{\thornado}{\texttt{thornado}}

\begin{document}

\title{Benchmarks of the Non-Relativistic Hydrodynamics Algorithms in \thornado}
%\author{Juliana Pulsinelli\altaffilmark{1} and Eirik Endeve\altaffilmark{2}}

%\altaffiltext{1}{Bearden High School, Knoxville, TN 37919-5449, USA}
%\altaffiltext{2}{Computational and Applied Mathematics Group, Oak Ridge National Laboratory, Oak Ridge, TN 37831-6354, USA; endevee@ornl.gov}

\maketitle

\begin{abstract}
\thornado\ is a new astrophysical code for radiation-hydrodynamics based on high-order Runge-Kutta Discontinuous Galerkin (RKDG) methods.  
It currently solves the Euler equations for fluid dynamics and the two-moment ($M_{1}$) approximation of the radiative transfer equation. 
In this document, we benchmark the algorithms for solving the Euler equations.  
An extensive set of one-dimensional Riemann problems with known exact solutions were run, and the \thornado\ results were graphed and compared to exact solutions in cases where they exist. 
\end{abstract}

\tableofcontents

\section{Introduction}

Astrophysical fluid dynamics models the movement of fluids in space. It yields very different solutions from terrestrial fluid dynamics, as the effect of gravity is reduced or removed. 
Fluid dynamics play a major role in the formation of stars and neutron stars, as well as the movement of galaxies and numerous other astrophysical topics.

\section{Equations and Numerical Method}

In this project, numerical algorithms for solving the Euler equations of fluid dynamics are tested against an extensive suite of standard problems taken from the literature.  
In the following subsections we briefly describe the equations and the numerical methods used.

\subsection{Euler Equations of Gas Dynamics}

The Euler equations for fluid dynamics \citep[see, e.g.,][for details]{Leveque2002} consists of the mass equation
\begin{equation}
  \frac{\partial \rho}{\partial t}  
  + \frac{\partial}{\partial x} \Big(\rho u\Big) = 0,
  \label{eq:massConservation}
\end{equation}
the momentum equation
\begin{equation}
  \frac{\partial}{\partial t} \Big(\rho u\Big)
  + \frac{\partial}{\partial x} \Big(\rho u^2 + p\Big) = 0,
  \label{eq:momentumConservation}
\end{equation}
and the energy equation
\begin{equation}
  \frac{\partial E}{\partial t}
  + \frac{\partial}{\partial x} \Big((E + p) u\Big) = 0.  
  \label{eq:energyConservation}
\end{equation}
where $\rho$ represents mass density, $u$ the fluid velocity, $p$ the pressure of the fluid, and $E$ the energy.  
Equation~\eqref{eq:massConservation} describes conservation of mass, Equation~\eqref{eq:momentumConservation} describes conservation of momentum, and equation \eqref{eq:energyConservation} describes conservation of energy.
The equation of state
\begin{equation}
  p = (\gamma - 1)
  \Big(E - \frac{1}{2} \rho u^2 \Big)
  \label{eq:eos}
\end{equation}
closes the system of equations, and links the pressure to the quantities evolved in Equations \eqref{eq:massConservation}-\eqref{eq:energyConservation}. 
The ratio of specific heats, $\gamma$, is taken to be a constant.

By defining the vector of conserved variables $\boldsymbol{U}=(\rho,\rho u,E)^{T}$ and the vector of fluxes $\boldsymbol{F}(\boldsymbol{U})=(\rho u,\rho u^{2}+p,(E+p)u)^{T}$ we write the Euler equations as system of conservation laws
\begin{equation}
  \partial_{t}\boldsymbol{U}
  +\partial_{x}\boldsymbol{F}(\boldsymbol{U})
  =0.
\end{equation}

\subsection{Runge-Kutta Discontinuous Galerkin Methods}

In the Runge-Kutta Discontinuous Galerkin \citep[RKDG; cf.][]{cockburnShu_2001} method the solution to the Euler equations is approximated by a local polynomial expansion of the form
\begin{equation}
  \boldsymbol{U}(x,t)
  \approx\boldsymbol{U}_{h}(x,t)=\sum_{j=0}^{k}\boldsymbol{C}_{j}(t)\,\phi_{j}(x),
\end{equation}
where $\{\phi_{j}(x)\}_{j=0}^{k}$ are polynomials of degree at most $k$ and $\boldsymbol{C}_{j}$ are expansion the coefficients to be solved for.  
In \thornado, $\phi_{j}(x)$ are given by Lagrange polynomials.

The computational domain $D=\{x:x\in[x_{\mbox{\tiny L}},x_{\mbox{\tiny R}}]\}$ is subdivided into $N$ elements $I_{i}=\{x:x\in(x_{i-1/2},x_{i+1/2})\}$.  
The width of element $I_{i}$ is then $\Delta x_{i}=x_{i+1/2}-x_{i-1/2}$.  
The semi-discrete DG method is then to find $\boldsymbol{U}_{h}$ such that
\begin{equation}
  \partial_{t}\int_{I_{i}}\boldsymbol{U}_{h}\,\phi_{j}\,dx
  +\big(\widehat{\boldsymbol{F}}\,\phi_{j}|_{i+1/2}-\widehat{\boldsymbol{F}}\,\phi_{j}|_{i-1/2}\big)
  -\int_{I_{i}}\boldsymbol{F}(\boldsymbol{U}_{h})\,\partial_{x}\phi_{j}\,dx=0
  \label{eq:semidiscreteDG}
\end{equation}
for all $\phi_{j}$ and all $I_{i}\in D$.  
In Equation~\eqref{eq:semidiscreteDG}, $\widehat{\boldsymbol{F}}_{i\pm1/2}$ is a numerical flux approximating the flux across the interfaces connecting neighboring elements.  
In \thornado, we have implemented the Local Lax-Friedrichs (LLF), the Harten-Lax-van Leer (HLL), and the HLLC Riemann solvers, but in the numerical experiments presented here, we only use the HLLC solver.  
To complete the spatial discretization, we use Gaussian quadratures to evaluate the integrals in Eq.~\eqref{eq:semidiscreteDG}.  
During time-integration, we apply limiters to the polynomial representation.  
We use the TVB limiter of \citet{} to prevent oscillations near discontinuities and the positivity limiter of \citet{ZhangShu2010} to ensure that the density and pressure remain positive.

\section{Numerical Experiments}

The code was tested using Riemann problems. The Riemann problem models a fluid in an area of one-dimensional space. The initial data features one jump discontinuity, and is constant on either side of this discontinuity. The computational domain is divided into space and time cells. The conditions of each cell are advanced forward for all space cells, one time step at a time, according to the Euler equations for fluid dynamics. \citep{Leveque2002}

%models an area with two fluids, each of which has different characteristics, which are separated from one another by a barrier. When the barrier is removed, the Riemann problem solver code models how the fluids mix, determining how different properties of the fluid change at various spatial locations over time. The fields are evolved forward over time based on the Euler equations.

Several tests were used to assess the accuracy of the integration method \citep{LiskaWendroff2003}. 
\begin{table}[h!]
\begin{center}
\begin{tabular}{ |c|c c c|c c c|c c| }
 \hline
 Test & $\rho_L$ & $u_L$ & $p_L$ & $\rho_R$ & $u_R$ & $p_R$ & $x_0$ & T \\
 \hline
 1 & 1 & 0 & 1 & 0.125 & 0 & 0.1 & 0.5 & 0.2 \\
 2 & 1 & 0.75 & 1 & 0.125 & 0 & 0.1 & 0.3 & 0.2 \\
 3 & 1 & -2 & 0.4 & 1 & 2 & 0.4 & 0.5 & 0.15 \\
 4 & 1 & 1 & $10^{-6}$ & 1 & -1 & $10^{-6}$ & 0.5 & 1 \\
 5 & 1 & -19.59745 & 1000 & 1 & -19.59745 & 0.01 & 0.8 & 0.012 \\
 6 & 5.99924 & 19.5975 & 460.894 & 5.99242 & -6.19633 & 46.095 & 0.4 & 0.035 \\
 7 & 1.4 & 0 & 1 & 1 & 0 & 1 & 0.5 & 2 \\
 8 & 1.4 & 0.1 & 1 & 1 & 0.1 & 1 & 0.5 & 2 \\
 9 & 0.1261192 & 8.9047029 & 782.92899 & 6.591493 & 2.2654207 & 3.1544874 & 0.5 & 0.0039 \\
10 & 5.6698 & 9.0299 & 100 & 1.0 & 0.0 & 1.0 & 0 & 1 \\
11 & 5.6698 & −1.4701 & 100 & 1.0 & -10.5 & 1.0 & 0 & 1 \\
 \hline
\end{tabular}
  \caption{table of tests}
  \label{tab:testtable}
\end{center}
\end{table}
$x_0$ refers to the position of the barrier. The computational domain is 0.0 to 1.0 in tests 1--8, 0.1 to 0.6 in test 9 \citep{LiskaWendroff2003}, -5 to 15 in test 10, and -15 to 5 in test 11 \citep{Leveque2002}.

Test 12 is an interacting blast waves test, computed on 0.0 to 1.0, with discontinuities at 0.1 and 0.9. At all positions, initial density is 1 and initial velocity is 0. Initial pressure is 1000 on the left, 0.01 in the center, and 100 on the right \citep{LiskaWendroff2003}.
%\clearpage
\subsection{Test 1: Sod's Shock Tube}

First, the tests were run with values computed at different intervals; the method of integration is the same in each test. Between the two endpoints, 128, 256, 512, and 1024 points were computed for on each respective test. Figures~\ref{fig:dencellcomp_20}--\ref{fig:pvcellcomp} show results of these tests:
\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=.95\textwidth]{dencellcomp_20}
  \end{center}
  \caption{Results showing the density for Sod's shock tube at $t=0.2$, computed with the first-order method using different number of elements: 128 (solid red), 256 (dashed green), 512 (dotted blue), and 1024 (dash-dot pink).  The exact solution, obtained with Toro's Riemann solver, is given by the dash-dotted cyan line.}
  \label{fig:dencellcomp_20}
\end{figure}
%\newpage

\begin{figure}
  \begin{center}
    \begin{tabular}{cc}
      \includegraphics[width=.475\textwidth]{velcellcomp_20} &
      \includegraphics[width=.475\textwidth]{prscellcomp_20}
    \end{tabular}
  \end{center}
  \caption{Results from the same experiments as shown in Figure~\ref{fig:dencellcomp_20}:
  velocity (left panel) and pressure (right panel).}
  \label{fig:pvcellcomp}
\end{figure}
%\newpage

%\clearpage
%\begin{figure}
 % \begin{center}

 % \end{center}
 % \caption{velocity}
 % \label{fig:velcellcomp_20}
%\end{figure}
%\clearpage
Next, several tests were run of different integration methods; the number of points was held constant at 128 each time. Higher numbers in the key correspond to higher-order, more exact methods of integration.

\begin{figure}[h]
  \begin{center}
    \includegraphics[width=.78\textwidth]{den128comp_20}
  \end{center}
  \caption{Results showing the density for Sod's shock tube at $t=0.2$, computed with 128 cells with methods of different orders: 1st (solid red), 2nd (dashed green) and 3rd (dotted blue). The exact solution, obtained with Toro's Riemann solver, is given by the dash-dotted pink line.}
  \label{fig:den128comp}
\end{figure}
%\newpage

\begin{figure}
  \begin{center}
	\begin{tabular}{cc}
      \includegraphics[width=.425\textwidth]{vel128comp_20} &
	  \includegraphics[width=.425\textwidth]{prs128comp_20}
	\end{tabular}
  \end{center}
  \caption{Results from the same experiments as shown in Figure~\ref{fig:den128comp}:
  velocity (left panel) and pressure (right panel).}
\end{figure}

These graphs show that the combination of both the lowest order integration method and 128 points yields significant error; however, improving either the method of integration or the number of points gives a solution much closer to the exact solution.

\clearpage

\subsection{Test 2}
For subsequent tests, the number of cells was consistent within each test, and the integration method was varied as in the first test. 
The graphs for the second test show a result common throughout many of the different tests: The first order integration method is considerably far from the exact solution; the second and third order integration methods yield nearly identical results which are near, but not exactly equal to, the exact solution. 
\begin{figure}[h]
  \begin{center}
     \includegraphics[width=.78\textwidth]{den_T2.png}	
  \end{center}
  \caption{Results showing the density for test 2 at $t=0.2$, computed with 100 cells with methods of different orders: 1st (solid red), 2nd (dashed green) and 3rd (dotted blue). The exact solution, obtained with Toro's Riemann solver, is given by the dash-dotted pink line.}
  \label{fig:den_T2}
\end{figure}

\begin{figure}
  \begin{center}
	\begin{tabular}{cc}
      \includegraphics[width=.425\textwidth]{vel_T2.png} &
	  \includegraphics[width=.425\textwidth]{prs_T2.png}
	\end{tabular}
  \end{center}
  \caption{Results from the same experiments as shown in Figure~\ref{fig:den_T2}:
  velocity (left panel) and pressure (right panel).}
\end{figure}

\clearpage

\subsection{Test 3}
Test 3 shows results similar to Test 2. The deviation of experimental results from exact is especially noticeable in velocity. Near the initial point of discontinuity, the exact results appear completely horizontal. Method 1 almost completely ignores this horizontal part, while Methods 2 and 3 smooth it out considerably. 
\begin{figure}[h]
  \begin{center}
     \includegraphics[width=.78\textwidth]{den_T3.png}	
  \end{center}
  \caption{Results showing the density for test 3 at $t=0.15$, computed with 100 cells with methods of different orders: 1st (solid red), 2nd (dashed green) and 3rd (dotted blue). The exact solution, obtained with Toro's Riemann solver, is given by the dash-dotted pink line.}
  \label{fig:den_T3}
\end{figure}

\begin{figure}
  \begin{center}
	\begin{tabular}{cc}
      \includegraphics[width=.425\textwidth]{vel_T3.png} &
	  \includegraphics[width=.425\textwidth]{prs_T3.png}
	\end{tabular}
  \end{center}
  \caption{Results from the same experiments as shown in Figure~\ref{fig:den_T3}:
  velocity (left panel) and pressure (right panel).}
\end{figure}

\clearpage

\subsection{Test 4}

This test obviously displays some difference from the exact solution in density. A depression appears in the center of the graph which is not part of the exact solution. This depression is most severe when method 1 is used, but it also appears with methods 2 and 3.

\begin{figure}[h]
  \begin{center}
     \includegraphics[width=.78\textwidth]{den_T4.png}	
  \end{center}
  \caption{Results showing the density for the Noh problem at $t=1$, computed with 100 cells with methods of different orders: 1st (solid red), 2nd (dashed green) and 3rd (dotted blue). The exact solution, obtained with Toro's Riemann solver, is given by the dash-dotted pink line.}
  \label{fig:den_T4}
\end{figure}

\begin{figure}
  \begin{center}
	\begin{tabular}{cc}
      \includegraphics[width=.425\textwidth]{vel_T4.png} &
	  \includegraphics[width=.425\textwidth]{prs_T4.png}
	\end{tabular}
  \end{center}
  \caption{Results from the same experiments as shown in Figure~\ref{fig:den_T4}:
  velocity (left panel) and pressure (right panel).}
\end{figure}

\clearpage

\subsection{Test 5}
This test has interesting features in velocity. The first integration method underestimates the velocity at positions between about 0.2 and 0.5. The second and third order methods find a bump in density at around 0.4, which is not present in the exact solution.

\begin{figure}[h]
  \begin{center}
     \includegraphics[width=.78\textwidth]{den_T5.png}	
  \end{center}
  \caption{Results showing the density for test 5 at $t=0.012$, computed with 200 cells with methods of different orders: 1st (solid red), 2nd (dashed green) and 3rd (dotted blue). The exact solution, obtained with Toro's Riemann solver, is given by the dash-dotted pink line.}
  \label{fig:den_T5}
\end{figure}

\begin{figure}
  \begin{center}
	\begin{tabular}{cc}
      \includegraphics[width=.425\textwidth]{vel_T5.png} &
	  \includegraphics[width=.425\textwidth]{prs_T5.png}
	\end{tabular}
  \end{center}
  \caption{Results from the same experiments as shown in Figure~\ref{fig:den_T5}:
  velocity (left panel) and pressure (right panel).}
\end{figure}
\clearpage

\subsection{Test 6}
For this test's pressures and velocity, all three integration methods give values very close to exact. However, method 1 proves to be quite bad--worse than usual--at computing the density for this test.

\begin{figure}[h]
  \begin{center}
     \includegraphics[width=.78\textwidth]{den_T6.png}	
  \end{center}
  \caption{Results showing the density for test 6 at $t=0.035$, computed with 200 cells with methods of different orders: 1st (solid red), 2nd (dashed green) and 3rd (dotted blue). The exact solution, obtained with Toro's Riemann solver, is given by the dash-dotted pink line.}
  \label{fig:den_T6}
\end{figure}

\begin{figure}
  \begin{center}
	\begin{tabular}{cc}
      \includegraphics[width=.425\textwidth]{vel_T6.png} &
	  \includegraphics[width=.425\textwidth]{prs_T6.png}
	\end{tabular}
  \end{center}
  \caption{Results from the same experiments as shown in Figure~\ref{fig:den_T6}:
  velocity (left panel) and pressure (right panel).}
\end{figure}

\clearpage

\subsection{Test 7}
This test features an unusual situation--method 1 calculates the velocity exactly, while method 2 introduces more error and method 3 introduces the most. However, this graph only appears extreme because the intervals on its axis are so small. The numbers are always very close to zero, but very slight errors occur in some of the computations.

\begin{figure}[h]
  \begin{center}
     \includegraphics[width=.78\textwidth]{den_T7.png}	
  \end{center}
  \caption{Results showing the density for test 7 at $t=2$, computed with 100 cells with methods of different orders: 1st (solid red), 2nd (dashed green) and 3rd (dotted blue). The exact solution, obtained with Toro's Riemann solver, is given by the dash-dotted pink line.}
  \label{fig:den_T7}
\end{figure}

\begin{figure}
  \begin{center}
	\begin{tabular}{cc}
      \includegraphics[width=.425\textwidth]{vel_T7.png} &
	  \includegraphics[width=.425\textwidth]{prs_T7.png}
	\end{tabular}
  \end{center}
  \caption{Results from the same experiments as shown in Figure~\ref{fig:den_T7}:
  velocity (left panel) and pressure (right panel).}
\end{figure}

\clearpage

\subsection{Test 8}

\begin{figure}[h]
  \begin{center}
     \includegraphics[width=.78\textwidth]{den_T8.png}	
  \end{center}
  \caption{Results showing the density for test 8 at $t=2$, computed with 100 cells with methods of different orders: 1st (solid red), 2nd (dashed green) and 3rd (dotted blue). The exact solution, obtained with Toro's Riemann solver, is given by the dash-dotted pink line.}
  \label{fig:den_T8}
\end{figure}

\begin{figure}
  \begin{center}
	\begin{tabular}{cc}
      \includegraphics[width=.425\textwidth]{vel_T8.png} &
	  \includegraphics[width=.425\textwidth]{prs_T8.png}
	\end{tabular}
  \end{center}
  \caption{Results from the same experiments as shown in Figure~\ref{fig:den_T8}:
  velocity (left panel) and pressure (right panel).}
\end{figure}

\clearpage

\subsection{Test 9}
When finding density, method 1 falls considerably short of the maximum. On the pressure graph, method 3 features both a jump above and a dip below the graph between positions 0.1 and 0.2. Velocity shows a similar (but considerably larger) dip, followed by jump, in the vertical section of the graph between 0.1 and 0.2. Method 1 also overestimates the maximum velocity and is above both exact and method 3 in the middle horizontal section.

\begin{figure}[h]
  \begin{center}
	\begin{tabular}{cc}
      \includegraphics[width=.4\textwidth]{den_T9.png} &
	  \includegraphics[width=.4\textwidth]{denT9zoom.png} \\
	  \includegraphics[width=.4\textwidth]{vel_T9.png} &	
      \includegraphics[width=.4\textwidth]{prs_T9.png}\\
	\end{tabular}	
  \end{center}
  \caption{Results for the peak problem at $t=0.0039$, computed with 800 cells with methods of different orders: 1st (solid red), 2nd (dashed green) and 3rd (dotted blue). The exact solution, obtained with Toro's Riemann solver, is given by the dash-dotted pink line. Density is shown in the top two panels at two different zoom levels. The bottom left panel shows velocity, and the bottom right panel shows pressure.}
\end{figure}

\clearpage

\subsection{Test 10}

\begin{figure}[h]
  \begin{center}
	\begin{tabular}{cc}
      \includegraphics[width=.4\textwidth]{den_T10.png} &
	  \includegraphics[width=.4\textwidth]{den10zoom.png} \\
	  \includegraphics[width=.4\textwidth]{vel_T10.png} &	
	  \includegraphics[width=.4\textwidth]{vel10zoom.png} \\
      \includegraphics[width=.4\textwidth]{prs_T10.png} &	
      \includegraphics[width=.4\textwidth]{prs10zoom.png} \\
	\end{tabular}	
  \end{center}
  \caption{Results for test 10 at $t=1$, computed with 400 cells with methods of different orders: 1st (solid red), 2nd (dashed green) and 3rd (dotted blue). The exact solution, obtained with Toro's Riemann solver, is given by the dash-dotted pink line. Each pair of side-by-side panels shows the same results at different zoom levels. The top panels show density, the middle show velocity, and the bottom show pressure.}
\end{figure}

\clearpage

\subsection{Test 11}

\begin{figure}[h]
  \begin{center}
	\begin{tabular}{cc}
      \includegraphics[width=.4\textwidth]{den_T11.png} &
	  \includegraphics[width=.4\textwidth]{den11zoom.png} \\
	  \includegraphics[width=.4\textwidth]{vel_T11.png} &	
	  \includegraphics[width=.4\textwidth]{vel11zoom.png} \\
      \includegraphics[width=.4\textwidth]{prs_T11.png} &	
      \includegraphics[width=.4\textwidth]{prs11zoom.png} \\
	\end{tabular}	
  \end{center}
  \caption{Results for test 11 at $t=1$, computed with 400 cells with methods of different orders: 1st (solid red), 2nd (dashed green) and 3rd (dotted blue). The exact solution, obtained with Toro's Riemann solver, is given by the dash-dotted pink line. Each pair of side-by-side panels shows the same results at different zoom levels. The top panels show density, the middle show velocity, and the bottom show pressure.}
\end{figure}

\clearpage

\subsection{Test 12: Interacting Blast Waves}
This test was run for all three integration methods with both 400 cells and 2000 cells. Since it has no exact solution, the 2000 cell, third order test is compared to all three 400 cell tests and labeled as ``exact.''

\begin{figure}[h]
  \begin{center}
     \includegraphics[width=.78\textwidth]{den_T12_400.png}	
  \end{center}
  \caption{Results showing the density for the Woodward-Collela blast wave problem at $t=1$, computed with 400 cells with methods of different orders: 1st (solid red), 2nd (dashed green) and 3rd (dotted blue). The solution labelled ``exact' and given by the dash-dotted pink line is the results of the problem when computed using the third-order integration method with 2000 cells.}
  \label{fig:den_T12_400}
\end{figure}

\begin{figure}[h]
  \begin{center}
     \includegraphics[width=.78\textwidth]{den_T12_2000.png}	
  \end{center}
  \caption{Results for the density from the same experiments as shown in Figure~\ref{fig:den_T12_400} computed with 2000 cells}
\end{figure}

\begin{figure}[h]
  \begin{center}
     \includegraphics[width=.78\textwidth]{vel_T12_400.png}	
  \end{center}
  \caption{Results for the velocity from the same experiments as shown in Figure~\ref{fig:den_T12_400}}
\end{figure}

\begin{figure}[h]
  \begin{center}
     \includegraphics[width=.78\textwidth]{vel_T12_2000.png}	
  \end{center}
  \caption{Results for the velocity from the same experiments as shown in Figure~\ref{fig:den_T12_400} computed with 2000 cells}
\end{figure}

\begin{figure}[h]
  \begin{center}
     \includegraphics[width=.78\textwidth]{prs_T12_400.png}	
  \end{center}
  \caption{Results for the pressure from the same experiments as shown in Figure~\ref{fig:den_T12_400}}
\end{figure}

\begin{figure}[h]
  \begin{center}
     \includegraphics[width=.78\textwidth]{prs_T12_2000.png}	
  \end{center}
  \caption{Results for the pressure from the same experiments as shown in Figure~\ref{fig:den_T12_400} computed with 2000 cells}
\end{figure}

\clearpage

\subsection{Test 13: Shock Entropy Waves}

This test was first run with 1000 cells to see the solution with high resolution. Subsequent tests all had 200 cells, with various values of the beta parameter. Again, the 3rd order, 1000 cell solution is called ``exact.''

\begin{figure}[h]
  \begin{center}
     \includegraphics[width=.78\textwidth]{1000.png}	
  \end{center}
  \caption{density, 1000 cells}
\end{figure}

\begin{figure}
  \begin{center}
	\begin{tabular}{cc}
      \includegraphics[width=.425\textwidth]{10.png} &
	  \includegraphics[width=.425\textwidth]{10zoom.png}
	\end{tabular}
  \end{center}
  \caption{density, beta=10}
  \label{fig:pvcellcomp}
\end{figure}

\begin{figure}
  \begin{center}
	\begin{tabular}{cc}
      \includegraphics[width=.425\textwidth]{50.png} &
	  \includegraphics[width=.425\textwidth]{50zoom.png}
	\end{tabular}
  \end{center}
  \caption{density, beta=50}
  \label{fig:pvcellcomp}
\end{figure}

\begin{figure}
  \begin{center}
	\begin{tabular}{cc}
      \includegraphics[width=.425\textwidth]{300.png} &
	  \includegraphics[width=.425\textwidth]{300zoom.png}
	\end{tabular}
  \end{center}
  \caption{density, beta=300}
  \label{fig:pvcellcomp}
\end{figure}

\bibliographystyle{apj}
\bibliography{references.bib}

\end{document}