### 2018-01-22 Documents/M1/realizableFermionicM1/realizableM1_fermionPaper.tex
Created tex files for the realiazableM1 paper.

\documentclass[review]{elsarticle}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

\journal{Journal of Computational Physics }

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{Realizability-Preserving Two-Moment DG Method for Fermionic Transport \tnoteref{ldrd-support}}
\tnotetext[ldrd-support]{This research is sponsored by ...}

%% Group authors per affiliation:
\author[ornl,utk-phys]{Eirik Endeve\corref{cor}}
\ead{endevee@ornl.gov}

\author[utk-phys]{Ran Chu}
\ead{rchu@vols.utk.edu}

\author[utk-phys]{Anthony Mezzacappa}
\ead{mezz@utk.edu}

\cortext[cor]{Corresponding author. Tel.:+1 865 576 6349; fax:+1 865 241 0381}

\address[ornl]{Computational and Applied Mathematics Group,
Oak Ridge National Laboratory,
Oak Ridge, TN 37831 USA }

\address[utk-phys]{Department of Physics and Astronomy,
University of Tennessee
Knoxville, TN 37996-1200}

\begin{abstract}
This paper ...
\end{abstract}

\begin{keyword}
Boltzmann equation, 
Radiation transport, 
Hyperbolic conservation laws, 
Discontinuous Galerkin, 
Realizability
\end{keyword}

\end{frontmatter}

\linenumbers

\input{sections/intro}
\input{sections/model}
\input{sections/MomentsRealizability}
\input{sections/IMEX}
\input{sections/DGmethod}
\input{sections/DGIMEXscheme}
\input{sections/limiter}
\input{sections/numerical}
\input{sections/conclusion}
%\input{sections/appendix}
\clearpage

\section*{References}

\bibliography{./refs/dgCurv.bib}
\end{document}