### 2018-08-14 Documents/M1/realizableFermionicM1/sections/conclusion.tex
Added the conclusion sectrion draft

\section{Conclusions}
\label{sec:conclusions}

We develop a realizability-preserving scheme which includes a suitable CFL condition, a realizability-preserving limiter, a realizability-preserving closure procedure and a realizability-preserving IMEX time integrator.
Thanks for the meaningful result domain being a convex set, the realizability-preservation of the limiter and the IMEX time integrator can be ensured by requiring they have convex combination forms.
And a realizability-preserving closure procedure can be obtained by any closure procedure building on Fermi-Dirac statistics.
In the physics which we interested, the transport system introduces a stiff problem: the collision term has much larger eigenvalues than that the advection term has.
Therefore, requiring the scheme be streaming accurate and diffusion accurate is an essential and achievable goal than being overall accurate.
We purpose our PD-ARS scheme. It is 2nd-order accurate and strong stability preserving in the streaming limit and works well in the diffusion limit. Numerical examples are provided to demonstrate those properties.

It's widely known that a realizability-preserving scheme is required for solving the approximating transport problems with conservation law, including the two-moment model.
However, few people are aware of the significance of realizability-preserving closure procedure: the Minerbo closure, which is not built on Fermi-Dirac distribution, is widely used.
We highly recommend the CB closure other than the Minerbo closure because the CB closure not only preserves the realizability but also requires no more calculation than the Minerbo closure.

The realizability-preservation of our IMEX scheme is universal: it can be implemented to any problems where the physical constraints confine its meaningful result to a convex set and the diffusion limit is important.

To build this work, we assume an easy collision term and no relativistic term on the left-hand side of the transport equation.
Hence, to solve a more realistic problem, a complex collision term and some relativistic terms should be taken into consideration.
Besides, a higher order accurate method is always desired.
Moreover, there is no reason to restrict the time integrator to be IMEX.
They all bring the remained open problems which require the future contribution.