### 2018-08-17 Documents/M1/realizableFermionicM1/sections/conclusion.tex
Modified the conclusion section.

\section{Conclusions}
\label{sec:conclusions}

We have developed a realizability-preserving DG-IMEX scheme for the two-moment method of fermion transport.
The scheme includes a suitable CFL condition, a realizability-preserving limiter, a realizable closure procedure and a realizability-preserving IMEX time integrator.
Its suitable CFL condition is limited by the phase space advection term only and can be as large as the CFL condition of the forward Euler time stepping scheme. 
The limiter enforces the spacial space discretization be constraint preserving for DG scheme.
The closure provides a realizable moment approximation.
And the IMEX time integrator, which can be written as a combination of forward Euler method and a backward Euler method, enforces each time step and each stage inside a time step be realizable.
Since the realizable domain is a convex set, the realizability-preserving property of the limiter and the IMEX time integrator can be ensured by requiring they have convex combination forms.
And a realizable closure procedure can be a closure procedure based on Fermi-Dirac statistics.

The collision term is stiff on time scale of interest: in a small domain, the transport is dominate by the diffusion; other than that, the streaming governed.
We tried first to build a formally second order IMEX scheme.
However, a correction step is needed which result a restricted CFL condition.
As an alternation, we separate the accuracy properties and require the scheme be second order in the streaming limit and first order in the diffusion limit.
This is an essential and more achievable goal.
Here we purpose our PD-ARS scheme. 
It is formally first order accurate but reduces to the optimal second-order strong-stability-preserving Runge-Kutta scheme in the streaming limit and works well in the diffusion limit. 
Numerical examples are provided to demonstrate those properties.

The constraint-preserving property of PD-ARS scheme is applicable to a wide class of problems where the physical constraints confine its meaningful result to a convex set and the diffusion limit is important.

It's known that a realizability-preserving scheme is required for solving the approximating transport problems with conservation law, including the two-moment model.
A break of the constraint-preserving property, such as implementing a non-realizable closure, the Minerbo closure for example, puts the conservation law in danger.
Among the simple realizable algebraic closures we discussed(CB, BL and Kershaw closure), we recommend the CB closure (reason ... ).

In this work, we adapted a relatively simple model. 
We assume a linear collision term with fixed background, only elastic scattering exists, and no relativistic term on the left-hand side of the transport equation.
Hence, to solve a more scientific problem, a complex collision term and some relativistic terms should be taken into consideration.
Besides, a higher order accurate method is always desired.
Moreover, there is no reason to restrict the time integrator to be IMEX.
In the context of constraint-preserving schemes, these extension provide additional challenge suitable for future investigations.