### 2018-10-22 Documents/M1/Astronum_2018/sections/Acknowledge.tex
Updated Astronum_2018.

\section{Acknowledgment} \label{se:Acknowledgment}

This research is sponsored, in part, by the Laboratory Directed Research and Development Program of Oak Ridge National Laboratory (ORNL), managed by UT-Battelle, LLC for the U. S. Department of Energy under Contract No. De-AC05-00OR22725.  
This research was supported by the Exascale Computing Project (17-SC-20-SC), a collaborative effort of the U.S. Department of Energy Office of Science and the National Nuclear Security Administration.  
This material is based, in part, upon work supported by the U.S. Department of Energy, Office of Science, Office of Advanced Scientific Computing Research.
Eirik Endeve and Anthony Mezzacappa acknowledge support from the National Science Foundation under group \rc{NSF-GP-... (number?)} and from the UT-ORNL Joint Institute for Computational Sciences.