### 2019-01-26 Documents/M1/Astronum_2018/sections/Conclusion.tex

\section{Conclusion}\label{se:Conclusion}

We have developed IMEX schemes suitable for a two-moment model of neutrino transport in \texttt{thornado} that obeys Fermi-Dirac statistics.
The scheme employs algebraic closure based on Fermi-Dirac statistics, high-order discontinuous Galerkin method for spatial discretization, and convex-invariant time integration to maintain realizability of the moments.  
Since the realizable domain is convex and its convexity can be inherited by a convex combination, some schemes having convex combinations as their stages should be able to preserve the realizable domain.
This idea encourages us to construct realizability-preserving time integrators, realizability-preserving IMEX schemes, and eventually a method with a realizability-preserving IMEX time integrator and high-order DG method.
See~\cite{chu_etal_2018} for more details and examples of the third-order DG methods.

In the applications that motivate this work, the neutrino distribution function can vary from 0 to 1.  
Hence, we have considered algebraic closures based on Fermi-Dirac statistics for both low and high occupancy.  
Among the seven algebraic closures we considered -- Kershaw~\cite{kershaw_1976}, Wilson~\cite{wilson_1975,leblancWilson_1970}, Levermore~\cite{levermore_1984}, Minerbo~\cite{minerbo_1978}, Janka 1~\cite{janka_1991}, Janka 2~\cite{janka_1992}, and Cernohorsky \& Bludman~\cite{cernohorskyBludman_1994} -- only the Cernohorsky \& Bludman closure obeys Fermi-Dirac statistics for all occupancies.  
As a result, we employed the Cernohorsky \& Bludman closure for the neutrino stationary state test in Section~\ref{se: Neutrino Stationary State Test}.
(We also ran the code with other algebraic closures for examination.
And the program crashed after a few time steps because of the nonphysical neutrino density/flux obtained and no correction step was taken for $f>1$.
We could make the code run by adding correction step, but the quality of the results may be compromised.
And it's difficult to determine whether the results are in fact compromised by a particular correction step and, if so, in what way and to what degree.)

Two PD-ARS schemes are proposed.
The one with SSPRK2 has second-order accuracy while the other with SSPRK3 has third-order accuracy, and both have the strong-stability preserving property in the streaming limit.  
Their accuracy was demonstrated on problems with known smooth solutions in streaming, absorption, and scattering-dominated regimes.
The neutrino transport test with emission, absorption, and isoenergetic scattering through a stationary background, was designed to test the convex-invariance of our PD-ARS schemes. 
The neutrino stationary state test shows that a method combining an algebraic closure based on Fermi-Dirac statistics and convex-invariant time integration can be a candidate for constructing a CCSN simulation code without "unphysical correction steps" that may break conservation laws.

In this work, we adopted Cartesian coordinates, a linear collision term, and a fixed non-relativistic material background.
More realistic problems of scientific interest, such as with energy-exchanging scattering and relativistic effects, are left for future research.